---
title: "Homework 1 - Sean"
output: html_document
date: "2024-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(igraph)
library(RColorBrewer)
library(tidyverse)


get.colors = function(cols, x) {
  ramp = colorRamp(cols)
  apply(ramp(x/max(x)), 1, function(y) rgb(y[1]/255, y[2]/255, y[3]/255))
}

```

Cleaning metadata into a table. DO NOT RUN. Read results from file.
```{r}
meta_data <- readLines("amazon-meta.txt.gz", encoding = 'utf-8')
meta_data <- meta_data[-1:-3]
print(meta_data[1:20])

stepi <- 0
pb <- txtProgressBar(min=0, max=length(meta_data), initial = 0, style = 3)



row <- 0 
meta_data_df <- data.frame(Id=NA, Title=NA,
                               Categories=NA, Group=NA,
                               Copurchased= NA, SalesRank=NA) 


Id <- ""
ASIN <- ""
Title <- ""
Categories <- ""
Group <- ""
Copurchased <- ""
SalesRank <- 0



start_time <- Sys.time()

for (i in meta_data){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)

  i <- trimws(i) 

  if (substr(i, 1 ,3) == "Id:"){
    Id <- trimws(substring(i ,4 ))
  }
  
  else if (substr(i, 1,5) == "ASIN:"){
    ASIN <- trimws(substring(i ,6 ))
  }
  
  else if (substr(i, 1,6) == "title:"){
    Title <- trimws(substring(i ,7 ))
  }
  
  else if (substr(i, 1,6) == "group:"){
    Group <- trimws(substring(i ,7 ))
  }
  
  else if (substr(i, 1,10) == "salesrank:"){
    SalesRank <- trimws(substring(i ,11 ))
  }
  
  else if (substr(i, 1,8) == "similar:"){
    Copurchased <- trimws(substring(i ,9 ))
  }
  
  else if (substr(i, 1,11) == "categories:"){
    Categories <- trimws(substring(i ,12 ))
  }
  
  else if (i == ""){
    
    row <- row + 1

      if (ASIN != ""){
        try(meta_data_df[row,'Id'] <- Id)
        try(meta_data_df[row,'Title'] <- Title)
        try(meta_data_df[row,'Categories'] <- Categories)
        try(meta_data_df[row,'Group'] <- Group)
        try(meta_data_df[row,'Copurchased'] <- Copurchased)
        try(meta_data_df[row,'SalesRank'] <- as.integer(SalesRank))
      }
      
    next
      
  Id <- ""
  ASIN <- ""
  Title <- ""
  Categories <- ""
  Group <- ""
  Copurchased <- ""
  SalesRank <- 0

    
  }
  
  
  
}

close(pb)

end_time <- Sys.time()
print(end_time - start_time)


print(meta_data_df)

write.csv(meta_data_df, 
          "C:\\Users\\Sean Law\\Desktop\\DBA4711\\Homework 1\\meta_data.csv",
          row.names=F)

```
















Data import
```{r}
df302 <- read.csv("amazon0302.txt.gz", header=TRUE, skip=3, sep="\t")
colnames(df302)[c(1,2)] <- c("from","to")
(df302)

df312 <- read.csv("amazon0312.txt.gz", header=TRUE, skip=3, sep="\t")
colnames(df312)[c(1,2)] <- c("from","to")
(df312)

df505 <- read.csv("amazon0505.txt.gz", header=TRUE, skip=3, sep="\t")
colnames(df505)[c(1,2)] <- c("from","to")
(df505)

df601 <- read.csv("amazon0601.txt.gz", header=TRUE, skip=3, sep="\t")
colnames(df601)[c(1,2)] <- c("from","to")
(df601)

```


Importing the already cleaned data that was previously cleaned above to save time
```{r}

meta_data_df <- read.csv("meta_data.csv")
meta_data_df

```



















### Actual work

Finding the degree centrality (to delete)
```{r}
G = graph_from_data_frame(df312, directed = T)
set.seed(45)
comp = components(G)
in.max.comp = (comp$membership == 1)
sg = induced_subgraph(G, in.max.comp)


temp <- degree(sg)
temp_df <- data.frame(temp)
temp_df <- cbind(as.integer(names(temp)), temp_df)
colnames(temp_df)[c(1,2)] <- c("Id","degree")


temp_df <- temp_df[temp_df$degree >= 30,]
temp_df

```

Getting degree centrality for each data set (Correct)

The reason why we are pruning for nodes with such high degree centrality is because it is 
unlikely that a node with low degree centrality will have high closeness centrality 
```{r}
get_degree_df <- function(data, pruning_value){
  #Function that return a dataframe containing node ID with their corresponding
  #degrees
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Reducing to the largest connected componenet of the graph
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)

  #Calculating degree
  temp <- degree(sg)
  
  #Converting into dataframe
  temp_df <- data.frame(temp)
  
  #Adding column with the node ID
  temp_df <- cbind(as.integer(names(temp)), temp_df)
  
  #Renaming columns
  colnames(temp_df)[c(1,2)] <- c("Id","degree")
  
  #Pruning 
  temp_df <- temp_df[temp_df$degree >= pruning_value,]
  temp_df <- temp_df[order(temp_df$degree, decreasing = T),]
  return(temp_df) 
}



#---------------------------------------------------------------------------

data_list <- list(df302, df312, df505, df601)
prune_list <- list(10,30,30,30)


stepi <- 0
pb <- txtProgressBar(min=0, max=length(data_list), initial = 0, style = 3)

for (i in 1:length(data_list)){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)
  
  print(get_degree_df(data_list[i], prune_list[i]))
}

close(pb)

```


SHRINKING DATA SET BY PRUNING BASED ON THE DEGREE CENTRALITY (ie the number of 
edges that each vertex has)
```{r}
#---------------------------------------------------------------------------
#try to prune the data based on these 
#so we first get all the values of items that we want to keep 

item_lst_302 <- get_degree_df(df302, 20)[,1]
pruned_302_df <- filter(df302, from %in% item_lst_302)
print(pruned_302_df)

item_lst_312 <- get_degree_df(df312, 50)[,1]
pruned_312_df <- filter(df312, from %in% item_lst_312)
print(pruned_312_df)

item_lst_505 <- get_degree_df(df505, 50)[,1]
pruned_505_df <- filter(df505, from %in% item_lst_505)
print(pruned_505_df)

item_lst_601 <- get_degree_df(df601, 50)[,1]
pruned_601_df <- filter(df601, from  %in% item_lst_601)
print(pruned_601_df)

```


Getting closeness centrality using the above pruned data
```{r}
get_closenees_centrality <- function(data){
  set.seed(45)
  G <- graph_from_data_frame(data, directed = T)
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)
  
  sg_cl <- closeness(sg)
  print(summary(sg_cl))
  sg_cl_df <- data.frame(sg_cl)
  sg_cl_df <- cbind(names(sg_cl), sg_cl_df)
  colnames(sg_cl_df)[c(1,2)] <- c("ID","closeness")
  sg_cl_df <- sg_cl_df[order(sg_cl_df$closeness, decreasing = T),]
  
  return(sg_cl_df) 
}


#---------------------------------------------------------------------------

data_list <- list(pruned_302_df, pruned_312_df, pruned_505_df, pruned_601_df)

stepi <- 0
pb <- txtProgressBar(min=0, max=length(data_list), initial = 0, style = 3)

for (i in 1:length(data_list)){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)
  
  print(get_closenees_centrality(data_list[i]))
}

close(pb)


```


Closeness centrality without pruning data
```{r}
data_list <- list(df302, df312, df505, df601)

stepi <- 0
pb <- txtProgressBar(min=0, max=length(data_list), initial = 0, style = 3)

for (i in 1:length(data_list)){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)
  
  print(get_closenees_centrality(data_list[i]))
}
close(pb)



#---------------------------------------------------------------------------
# TESTING
start_time <- Sys.time()
print(get_closenees_centrality(df302))
end_time <- Sys.time()
print(end_time - start_time)


```



Getting between-ness centrality with pruned data
```{r}
get_between_centrality <- function(data){
  set.seed(45)
  G <- graph_from_data_frame(data, directed = T)

  bn <- betweenness(G)
  print(summary(bn))
  bn_df <- data.frame(bn)
  bn_df <- cbind(names(bn), bn_df)
  colnames(bn_df)[c(1,2)] <- c("ID","between")
  bn_df <- bn_df[order(bn_df$between, decreasing = T),]
  
  return(bn_df) 
}


#---------------------------------------------------------------------------

data_list <- list(pruned_302_df, pruned_312_df, pruned_505_df, pruned_601_df)

stepi <- 0
pb <- txtProgressBar(min=0, max=length(data_list), initial = 0, style = 3)

for (i in 1:length(data_list)){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)
  
  print(get_closenees_centrality(data_list[i]))
}
close(pb)


```


Getting page-rank with pruned data
```{r}
get_page_rank <- function(data){
  set.seed(45)
  G <- graph_from_data_frame(data, directed = T)
  comp <- components(G)
  
  pr <- page.rank(G)$vector
  print(summary(pr))
  pr_df <- data.frame(pr)
  pr_df <- cbind(names(pr), pr_df)
  colnames(pr_df)[c(1,2)] <- c("ID","page_rank")
  pr_df <- pr_df[order(pr_df$between, decreasing = T),]
  
  return(pr_df) 
}


#---------------------------------------------------------------------------

data_list <- list(pruned_302_df, pruned_312_df, pruned_505_df, pruned_601_df)

stepi <- 0
pb <- txtProgressBar(min=0, max=length(data_list), initial = 0, style = 3)

for (i in 1:length(data_list)){
  
  stepi <- stepi + 1
  setTxtProgressBar(pb, stepi)
  
  print(get_closenees_centrality(data_list[i]))
}

close(pb)
```























### Functions that will be used

```{r}

top100_group <- function(data, top_, value_col){
  #this function takes in data, top_ and the value column and returns a dataframe containing the 
  #the top (top_) rows of data in each group for the value_col 
  out_df <- data %>% group_by(Group) %>%
    arrange(desc(value_col)) %>% 
    slice(1:top_)
  return(out_df)
}


```










### DATA WTIH METADATA FOR DEGREE CENTRALITY 
```{r}
get_degree_df_with_meta_data <- function(data){
  #Function that return a dataframe containing node ID with their corresponding
  #degrees
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Reducing to the largest connected componenet of the graph
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)

  #Calculating degree
  temp <- degree(sg)
  
  #Converting into dataframe
  temp_df <- data.frame(temp)
  
  #Adding column with the node ID
  temp_df <- cbind(as.integer(names(temp)), temp_df)
  
  #Renaming columns
  colnames(temp_df)[c(1,2)] <- c("Id","degree")
  
  #Reordering based on degree
  temp_df <- temp_df[order(temp_df$degree, decreasing = T),]
  
  #Left join meta data
  temp_df <- left_join(temp_df, meta_data_df, by= "Id")
  

  return(temp_df) 
}



#---------------------------------------------------------------------------
#Apply function 

df302_degree_meta <- get_degree_df_with_meta_data(df302)
print(df302_degree_meta)

df312_degree_meta <- get_degree_df_with_meta_data(df312)
print(df312_degree_meta)

df505_degree_meta <- get_degree_df_with_meta_data(df505)
print(df505_degree_meta)

df601_degree_meta <- get_degree_df_with_meta_data(df601)
print(df601_degree_meta)
```

TEST PROBLEM THERE ARE NOT 100 products in each group?!?!?! Ialready never remove any data
```{r}
get_degree_df_with_meta_data <- function(data){
  #Function that return a dataframe containing node ID with their corresponding
  #degrees
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Calculating degree
  temp <- degree(G)
  
  #Converting into dataframe
  temp_df <- data.frame(temp)
  
  #Adding column with the node ID
  temp_df <- cbind(as.integer(names(temp)), temp_df)
  
  #Renaming columns
  colnames(temp_df)[c(1,2)] <- c("Id","degree")
  
  #Reordering based on degree
  temp_df <- temp_df[order(temp_df$degree, decreasing = T),]
  
  #Left join meta data
  temp_df <- left_join(temp_df, meta_data_df, by= "Id")
  

  return(temp_df) 
}



#---------------------------------------------------------------------------

df302_degree_meta <- get_degree_df_with_meta_data(df302)
print(df302_degree_meta)

df312_degree_meta <- get_degree_df_with_meta_data(df312)
print(df312_degree_meta)

df505_degree_meta <- get_degree_df_with_meta_data(df505)
print(df505_degree_meta)

df601_degree_meta <- get_degree_df_with_meta_data(df601)
print(df601_degree_meta)
```



Group by group and then just head 100 
```{r}

top100_group(df302_degree_meta, 5, 'degree')
top100_group(df312_degree_meta, 5, 'degree')
top100_group(df505_degree_meta, 5, 'degree')
top100_group(df601_degree_meta, 5, 'degree')

```












### Getting closeness centrality 
```{r}
get_closeness_df_with_meta_data <- function(data){
  #Function that return a dataframe containing node ID with their corresponding
  #closeness
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Reducing to the largest connected componenet of the graph
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)

  #Calculating closeness
  sg_cl <- closeness(sg)
  print(summary(sg_cl))
  
  #Converting into dataframe
  sg_cl_df <- data.frame(sg_cl)
  
  #Adding column with the node ID
  sg_cl_df <- cbind(as.integer(names(sg_cl)), sg_cl_df)
  
  #Renaming columns
  colnames(sg_cl_df)[c(1,2)] <- c("Id","closeness")
  
  #Reordering based on degree
  sg_cl_df <- sg_cl_df[order(sg_cl_df$closeness, decreasing = T),]
  
  #Left join meta data
  sg_cl_df <- left_join(sg_cl_df, meta_data_df, by= "Id")
  

  return(sg_cl_df) 
}



#---------------------------------------------------------------------------
#Apply function 

df302_close_meta <- get_closeness_df_with_meta_data(pruned_302_df)
print(df302_close_meta)

df312_close_meta <- get_closeness_df_with_meta_data(pruned_312_df)
print(df312_close_meta)

df505_close_meta <- get_closeness_df_with_meta_data(pruned_505_df)
print(df505_close_meta)

df601_close_meta <- get_closeness_df_with_meta_data(pruned_601_df)
print(df601_close_meta)
```


Group by group and then just head 100 
```{r}

top100_group(df302_close_meta, 5, 'degree')
top100_group(df312_close_meta, 5, 'degree')
top100_group(df505_close_meta, 5, 'degree')
top100_group(df601_close_meta, 5, 'degree')

```













### Getting betweenness centrality 
```{r}
get_between_df_with_meta_data <- function(data){
  #Function that return a dataframe containing node ID with their corresponding
  #betweenness
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Reducing to the largest connected componenet of the graph
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)

  #Calculating betweenness
  bn <- betweenness(G)
  print(summary(bn))
  
  #Converting into dataframe
  bn_df <- data.frame(bn)
  
  #Adding column with the node ID
  bn_df <- cbind(as.integer(names(bn)), bn_df)
  
  #Renaming columns
  colnames(bn_df)[c(1,2)] <- c("Id","betweenness")
  
  #Reordering based on degree
  bn_df <- bn_df[order(bn_df$betweenness, decreasing = T),]
  
  #Left join meta data
  bn_df <- left_join(bn_df, meta_data_df, by= "Id")
  

  return(bn_df) 
}



#---------------------------------------------------------------------------
#Apply function 

df302_bet_meta <- get_between_df_with_meta_data(pruned_302_df)
print(df302_bet_meta)

df312_bet_meta <- get_between_df_with_meta_data(pruned_312_df)
print(df312_bet_meta)

df505_bet_meta <- get_between_df_with_meta_data(pruned_505_df)
print(df505_bet_meta)

df601_bet_meta <- get_between_df_with_meta_data(pruned_601_df)
print(df601_bet_meta)
```


Group by group and then just head 100 
```{r}

top100_group(df302_bet_meta, 5, 'degree')
top100_group(df312_bet_meta, 5, 'degree')
top100_group(df505_bet_meta, 5, 'degree')
top100_group(df601_bet_meta, 5, 'degree')

```
















### Getting page-rank centrality 
```{r}
get_page_rank_df_with_meta_data <- function(data){
  #Function that return a dataframe containing node ID with their corresponding
  #page-rank
  
  #creating igraph object using edge data and meta data
  G <- graph_from_data_frame(data, directed = T)
  
  #Reducing to the largest connected componenet of the graph
  comp <- components(G)
  in.max.comp <- (comp$membership == 1)
  sg <- induced_subgraph(G, in.max.comp)

  #Calculating betweenness
  pr <- page.rank(G)$vector
  print(summary(pr))
  
  #Converting into dataframe
  pr_df <- data.frame(pr)
  
  #Adding column with the node ID
  pr_df <- cbind(as.integer(names(pr)), pr_df)
  
  #Renaming columns
  colnames(pr_df)[c(1,2)] <- c("Id","page_rank")
  
  #Reordering based on degree
  pr_df  <- pr_df[order(pr_df$page_rank, decreasing = T),]
  
  #Left join meta data
  pr_df <- left_join(pr_df, meta_data_df, by= "Id")
  

  return(pr_df) 
}



#---------------------------------------------------------------------------
#Apply function 

df302_page_rank_meta <- get_page_rank_df_with_meta_data(pruned_302_df)
print(df302_bet_meta)

df312_page_rank_meta <- get_page_rank_df_with_meta_data(pruned_312_df)
print(df312_bet_meta)

df505_page_rank_meta <- get_page_rank_df_with_meta_data(pruned_505_df)
print(df505_bet_meta)

df601_page_rank_meta <- get_page_rank_df_with_meta_data(pruned_601_df)
print(df601_bet_meta)
```


Group by group and then just head 100 
```{r}

top100_group(df302_page_rank_meta, 5, 'degree')
top100_group(df312_page_rank_meta, 5, 'degree')
top100_group(df505_page_rank_meta, 5, 'degree')
top100_group(df601_page_rank_meta, 5, 'degree')

```


















